---
title: "Data Visualization - Basic Functions 01"
subtitle: "Exploring datasets"
author: "Create by Joniel B Barreto"
date: "2022-08-16"
output:
  html_document:
    df_print: paged
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
  html_notebook: default
---

## Install packages
```{r}
install.packages("dplyr")         # Install package
install.packages("data.table")
install.packages("matrixStats")
library(dplyr)                    # load the package
library(data.table)
library(matrixStats)
```

## Reading the dataset
The "drinks.txt" data shows information about alcohol consumption in 193 countries. It has 5 columns: country (chr), beer_servings (int), spirit_servings (int), wine_servings (int) and total_litres_of_pure_alcohol (dbl)
```{r}
df <- fread("drinks.txt", header = TRUE, sep = ",")
df
```
The above R code, assumes that the file “drinks.txt” is in your current working directory. To know your current working directory, type the function getwd() in R console. To set the directory, type the function setwd() in R console.

fread() function is a faster way to read files, particularly large files. The good thing about this function is that it automatically detects column types and separators, which can also be specified manually. It can also detect the header names and assign them to corresponding columns.

There's some functions to read files, such as read.table or readlines and others, but as we want to work with this data as a data frame and the file has special characters fread it's the best choice.


#### 1. What is the mean and standard deviation of beer (beer_servings), liquor (spirit_servings), wine (wine_servings) and total alcohol (total_litres_of_pure_alcohol) consumption?
```{R}
# Calculate the mean
df_m <- colMeans(df[,-1])
df_m
```
```{R}
# Calculate standard deviations
df_std <- apply(df[,-1],2,sd)
df_m
```
As the first column is formed by country names, the averages and the standard deviations are calculated disregarding the first column [,-1].

The apply function can be used to find the standard deviations for all columns. Here 2 refers to the columns, to find the standard deviations for rows then we just need to replace this 2 with 1.

```{R}
# Average beer consumption
df_m["beer_servings"]
```
```{R}
# Max liquor consumption value
max(df[,"spirit_servings"])
```
```{R}
# Min wine consumption value
min(df[,"wine_servings"])
```
```{R}
# Standard deviation of alcohol consumption
df_std["total_litres_of_pure_alcohol"]
```
```{R}
# Some statistical measures
summary(df)
```
#### 2. What is the consumption of liquor in France?
```{R}
# Data on alcohol consumption in Brazil
filter(df,country =='Brazil')
```

```{r}
# Consumption of liquor in France
filter(df,country =='France')$spirit_servings
```
#### 3. What are the 5 countries where the most wine is consumed?
```{r}
df[order(-df$wine_servings)][1:5]$country
```
To sort a data frame in R, you can use the order() function. By default, sorting is ASCENDING. The minus sign it's used to indicate DESCENDING order.

The [1:5] indicate that we want only the first five values and $country to show only the countries names.

#### 4. In which countries beer consumption is above average?
```{r}
# Countries whose total alcohol consumption is greater than or equal to 12 liters per person
filter(df,total_litres_of_pure_alcohol >= 12)
```
```{r}
# df_m["beer_servings"] is the average beer consumption
filter(df,beer_servings > df_m["beer_servings"])$country
```
#### 5. Which countries in which total alcohol consumption is higher than Brazil, in order of lowest to highest?
```{r}
# Total alcohol consumption in Brazil
alcool_brasil <-filter(df,country == 'Brazil')$total_litres_of_pure_alcohol
alcool_brasil
```
```{r}
# Countries with higher consumption than Brazil
df_alcool <- filter(df,total_litres_of_pure_alcohol > alcool_brasil)
# Sort from lowest to highest
df_alcool[order(df_alcool$total_litres_of_pure_alcohol)]
df_alcool
```
